### 2019-08-23 lower.tex
Minor fixes
simpler LB proof, first draft

In this section we consider the Online Balanced Partition problem.
We show a lower bound of $\Omega(k \cdot \ell)$ that uses an input that has a perfect partition.
Therefore, this result is a lower bound for the general Online Balanced Partition problem, as well as a lower bound for Online Balanced Partition with perfect partition.
The latter complements an upper bound $O(k \cdot \ell)$ for Online Balanced Partition with perfect partition that we present in Section TODO.

@@

@@

  Now, we claim that given the input $I_1I_2$, every component-respecting DET 
  % (that does not terminate prematurely and never resides in a configuration with components split)
  must produce a sequence $\langle x^{(i)}\rangle$ of length $L = k\cdot \ell - (k-1) - 2 - (\ell-1) - 1$.
  Consider a request $(x^{(i)}, c(x^{(i)}))$.
  If in this configuration at least one singleton component is present (other than the components of $(x^{(i)}, c(x^{(i)}))$), then serving this request is possible (i.e., the perfect partition exists), and hence a singleton node appears as $x^{(i+1)}$ and the sequence continues.
  Otherwise, no perfect partition exists, as we have a component $\{a_1, \ldots, a_{k-1}\}$ of size $k-1$ and it can only be complemented by a singleton component.

>>>>

In this section we consider the Online Balanced Partition problem.
We show a lower bound of $\Omega(k \cdot \ell)$ that uses an input that has a perfect partition.
Therefore, this result is a lower bound for the general Online Balanced Partition problem, as well as a lower bound for Online Balanced Partition with perfect partition.
The latter complements an upper bound $O(k \cdot \ell)$ for Online Balanced Partition with perfect partition that we present in Section \ref{sec:upperbound}.

@@

\begin{proof}
A simpler proof: \\
We construct an instance of the problem with $\ell$ clusters 
$\set{ S_1, S_2,\dots , S_{\ell}}, |S_i|  = k$.
Let $I(C)$ denote the cluster where nodes in a component $C$ are located initially.
Consider any online algorithm ALG.
We construct the input sequence of requests for ALG as follows.
First,
we issue $k-2$ (internal) requests so that ALG form a component of $k-1$
nodes on clusters $S_1$.
Note that this does not cost ALG.
Let $x_0$  be the only single nodes left on $S_1$ and  $y_0 \in S_2$ be any single node on $S_2$.
Next,
we issue the request between $x_0$ and $y_0$.
Since this is an external request,
ALG joins the into one component and collocates them in some cluster other than $S_1$ and $S_2$.
For this,
ALG moves to a new configuration that replaces $x_0$ and $y_0$ with two other single nodes $x_1$ and $y_1$ respectively.
Next,
we issue a request between $x_1$ and the largest component $C$ s.t.~$I(C) = I(v)$.
ALG must collocate $x_1$ and $C$ on some cluster other than $S_1$ and
consequently replaces $x_1$.
We repeat issuing requests between the single node $x_i$ on $S_1$ and the largest component $C'$ s.t.~$I(C')=I(x_i)$,
 until there are only two single nodes left that  originate from the same cluster.
I.e.,
two single nodes $x^*, y^*,I(x^*) = I(y^*)$
that constitute the last pair of such nodes.
At this point there are at most $\ell+1$ single nodes left
since there must be at least two nodes with the same initial cluster in any subset of $\ell+1$
nodes.
Given this sequence of requests,
the optimal strategy is to swap $x_0$ and $y_0$ with $x^*$ and $y^*$ respectively.
Hence,
OPT pays for $4$ migrations and
ALG pays at least one for each node in the sequence $X := x_0, x_1,\dots$.
We exclude at most $(k-1) + ( \ell+1)$ nodes out of $k.\ell$ nodes,
therefore $|X| \geq k.\ell - k - \ell \in \Omega(k.\ell)$.
\end{proof}

@@

  Now, we claim that given the input $I_1I_2$, every component-respecting DET 
  % (that does not terminate prematurely and never resides in a configuration with components split)
  must produce a sequence $\langle x^{(i)}\rangle$ of length $L = k\cdot \ell - (k-1) - 2 - (\ell-1) - 1$.
  \maciek{ToDo: a sentence that explains each summand of $L$}
  \maciek{Name of $L$ should indicate a connection to $I_2$, e.g., maybe just $|I_2|$, and then $L-\ell-1$ from last paragraph should be related to $I_2'$}
  Consider a request $(x^{(i)}, c(x^{(i)}))$.
  If in this configuration at least one singleton component is present (other than the components of $(x^{(i)}, c(x^{(i)}))$), then serving this request is possible (i.e., the perfect partition exists), and hence a singleton node appears as $x^{(i+1)}$ and the sequence continues.
  Otherwise, no perfect partition exists, as we have a component $\{a_1, \ldots, a_{k-1}\}$ of size $k-1$ and it can only be complemented by a singleton component.

### 2019-08-23 upper.tex
simpler LB proof, first draft

\begin{algorithm}
	\renewcommand{\algorithmicrequire}{\textbf{Input:}}
	\renewcommand{\algorithmicensure}{\textbf{Output:}}
	\begin{algorithmic}[1]
		\Require 
		$k, \ell$,
		initial partition $P_I$,
		sequence of  edges $\sigma_1, \dots, \sigma_N$ covering all ground truth components
		\Ensure Final partition $P_F$ 
		\State for each node $v$ create a singleton component $C_v$ and add it to $\mathcal{C}$ \label{line:initcomponents}
		\State on communication request $\sigma_t=\{u,v\}$:
		\State Let $C_1 \ni u$ and $C_2 \ni v$ be the container components
		\If{$C_1 \neq C_2$}
		\State unite the two components into a single component $C'$ and
		$\mathcal{C} = (\mathcal{C}\setminus\set{C_1, C_2}) \cup ~\set{C'}$ \label{line:mergecomponents}
		\If{cluster$(C_1) \neq$ cluster$(C_2)$}		
		\State \textit{re-partition($k, \ell, P_I, \mathcal{C}$)} \label{line:rebalance} 
		\EndIf
		\EndIf
	\end{algorithmic}
	\caption{Perfect Partition Learner}
	\label{alg:ppl}
      \end{algorithm}

@@

@@

\begin{theorem}	\label{thm:upperbound}
	Algorithm \ref{alg:ppl} reaches the final partition while being $(2.k.\ell)$-competitive.
\end{theorem}
\begin{IEEEproof}
	The algorithm eventually reaches the final partition since it
	 evaluates all possible $\ell$-way partitioning of components on each external request,
	including the request that completes revealing of all ground truth components.
	There are at most $(k-1).\ell < k.\ell $ calls to \emph{re-partition()} and by Lemma \ref{lemma:rebalancecost} each costs at most $2.OPT$.
	The total cost is therefore at most $2.OPT.k.\ell$ which implies the competitive ratio.
\end{IEEEproof}

>>>>

\begin{algorithm}
	\renewcommand{\algorithmicrequire}{\textbf{Input:}}
	\renewcommand{\algorithmicensure}{\textbf{Output:}}
	\begin{algorithmic}[1]
		\Require 
		$k, \ell$,
		initial partition $P_I$,
		sequence of  edges $\sigma_1, \dots, \sigma_N$ 
		\Ensure Final partition $P_F$ 
		\State for each node $v$ create a singleton component $C_v$ and add it to $\mathcal{C}$ \label{line:initcomponents}
		\State on communication request $\sigma_t=\{u,v\}$:
		\State Let $C_1 \ni u$ and $C_2 \ni v$ be the container components
		\If{$C_1 \neq C_2$}
		\State unite the two components into a single component $C'$ and
		$\mathcal{C} = (\mathcal{C}\setminus\set{C_1, C_2}) \cup ~\set{C'}$ \label{line:mergecomponents}
		\If{cluster$(C_1) \neq$ cluster$(C_2)$}		
		\State \textit{re-partition($k, \ell, P_I, \mathcal{C}$)} \label{line:rebalance} 
		\EndIf
		\EndIf
	\end{algorithmic}
	\caption{Perfect Partition Learner}
	\label{alg:ppl}
      \end{algorithm}

@@

\mahmoud{address the \textit{re-partition()}}

@@

\begin{theorem}	\label{thm:upperbound}
	Algorithm \ref{alg:ppl} reaches the final partition while being $(2.k.\ell)$-competitive.
\end{theorem}
\begin{IEEEproof}
	The algorithm eventually reaches the final partition since it
	 evaluates all possible $\ell$-way partitions of components on each external request,
	including the request that completes revealing of all ground truth components.
	There are at most $(k-1).\ell < k.\ell $ calls to \emph{re-partition()} and by Lemma \ref{lemma:rebalancecost} each costs at most $2.OPT$.
	The total cost is therefore at most $2.OPT.k.\ell$ which implies the competitive ratio.
\end{IEEEproof}
